Offshore Modeling - Manuscript Model

### 2020-02-10 2019_edits/R/offshore_analyses/final_analyses/Manuscript_Offshore_Modeling.Rmd
made more changes, added link to index file

---
title: "Offshore Modeling - Manuscript Model"
author: "Adam A. Kemberling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

@@

```{r load export}
seamap <- read_csv(str_c("~/Dropbox/SEAMAP_2019/data/offshore/", 
                        "seamap_manuscriptmods_full.csv"), 
                   col_types = cols()) %>% 
  mutate(year_f = factor(year_f),
         Season = factor(season, levels = c("Summer", "Fall")),
         region = factor(region, levels = c("Texas", "Louisiana", "MS Bight", "Florida"))
)

seamap_west <- read_csv(str_c("~/Dropbox/SEAMAP_2019/data/offshore/", 
                        "seamapwest_manuscriptmods_full.csv"), 
                        col_types = cols()) %>% 
  mutate(year_f = factor(year_f),
         Season = factor(season, levels = c("Summer", "Fall")),
         region = factor(region, levels = c("Texas", "Louisiana", "MS Bight", "Florida"))
)
```

@@

## Yearly Catch Plot {.tabset}

@@

*NOTE: Contains Data Across All Years and Regions (Including Florida)*

```{r}
year_sums <- seamap %>% 
  group_by(survey_year) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(cpue_towspd)),
            `Mean Catch`      = mean(cpue_towspd, na.rm = T),
            `SD`              = sd(cpue_towspd, na.rm = T),
            `Var`             = var(cpue_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(crab_presence, na.rm = T),
            `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)
            ) %>% 
  ungroup() %>% 
    mutate(Survey_Year        = factor(survey_year),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
         )

@@

year_plot <- year_sums %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot() +
    geom_col(aes(survey_year, `Mean Catch`)) +
    geom_errorbar(aes(x = survey_year, 
                      ymin = catch_ymin,
                      ymax = catch_ymax),
                  size = .25) +
  labs(x = NULL,
       y = "Mean Catch (C. sapidus / Hectare)")

@@

```{r}
year_foa <- year_sums %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot() +
    geom_col(aes(survey_year, `% Occurrence`)) +
    # geom_errorbar(aes(x = survey_year, 
    #                   ymin = occurrence_ymin,
    #                   ymax = occurrence_ymax),
    #               size = .25) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL, y = "Percent Occurrence")

@@

@@

year_plot + year_foa
```

## Region - Season Plot {.tabset}

*NOTE: Contains Data Across All Years and Regions (Including Florida)*

*NOTE 2: Mean Catch (per hectare) and frequency of occurrence calculated for each year, in each region, for each season. The mean and standard error of those annual measurements are then taken*

```{r }
# rs_summary <- seamap %>% 
#   filter(is.na(region)      == FALSE) %>% 
#   group_by(region, season) %>% 
#   summarise(Stations          = n(),
#             `Missing Obs`     = sum(is.na(cpue_towspd)),
#             `Mean Catch`      = mean(cpue_towspd, na.rm = T),
#             `SD`              = sd(cpue_towspd, na.rm = T),
#             `Var`             = var(cpue_towspd),
#             `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
#             `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
#             `% Occurrence`    = mean(crab_presence, na.rm = T),
#             `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
#     ungroup() %>% 
#     mutate(Season             = factor(season, levels = c("Summer", "Fall")),
#            region             = factor(region, 
#                                        levels = c("Texas", "Louisiana", "MS Bight", "Florida")),
#            #region             = fct_rev(region),
#            catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
#            catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
#            catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
#            occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
#            occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
#            occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
#     )

# #Summary Table
# rs_summary %>%
#   select(Region = region, Season, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>%
#   knitr::kable()

#Calaculate Annual average cpue and percent occurrence
rs_summary <- seamap %>% 
  group_by(survey_year, region, season) %>% 
  summarise(cpue = mean(cpue_towspd, na.rm = T),
            percent_occurrence = mean(crab_presence, na.rm = T)) %>% 
  group_by(region, season) %>% 
  summarise(nyears = n(),
            avg_annual_cpue = mean(cpue, na.rm = T),
            annual_cpue_se = sd(cpue, na.rm = T) / sqrt(nyears),
            avg_percent_occurrence = mean(percent_occurrence, na.rm = T),
            po_se = sd(percent_occurrence, na.rm = T) / sqrt(nyears)) %>% 
    ungroup() %>%
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           region             = factor(region, levels = c("Texas", "Louisiana", 
                                                          "MS Bight", "Florida")),
           catch_ymin         = avg_annual_cpue - (2 * annual_cpue_se),
           catch_ymax         = avg_annual_cpue + (2 * annual_cpue_se),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = avg_percent_occurrence - (2 * po_se),
           occurrence_ymax    = avg_percent_occurrence + (2 * po_se),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin),
           occurrence_ymax    = ifelse(occurrence_ymax > 1, 1, occurrence_ymax)
    )

@@

# Plot
rs_plot <- rs_summary %>% 
  ggplot(aes(region, avg_annual_cpue, fill = season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  position = "dodge", 
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = NULL,
       y = "Mean Catch (C. sapidus / Hectare)") + 
  theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

```{r}
rs_plot_2 <- rs_summary %>% 
  ggplot(aes(region, avg_percent_occurrence, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL,
       y = "Percent Occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

```{r}
rs_plot + rs_plot_2

@@

## Seasonal Catch Plot {.tabset}

*NOTE 2: Mean Catch (per hectare) and frequency of occurrence calculated for each year, for each season. The mean and standard error of those annual measurements are then taken*

### Season Factor Summary

```{r}
# season_summary <- seamap %>% 
#   group_by(survey_year, season) %>% 
#   summarise(Stations          = n(),
#             `Missing Obs`     = sum(is.na(cpue_towspd)),
#             `Mean Catch`      = mean(cpue_towspd, na.rm = T),
#             `SD`              = sd(cpue_towspd, na.rm = T),
#             `Var`             = var(cpue_towspd),
#             `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
#             `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
#             `% Occurrence`    = mean(crab_presence, na.rm = T),
#             `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
#     ungroup() %>% 
#     mutate(survey_year        = factor(survey_year),
#            # StatZone         = factor(StatZone),
#            # StatZone         = fct_rev(StatZone),
#            catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
#            catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
#            catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
#            occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
#            occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
#            occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
#          )
# 
# season_summary %>% select(`Survey Year` = survey_year, Season = season, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>% knitr::kable()

#Calaculate Annual average cpue and percent occurrence
season_summary <- seamap %>% 
  group_by(survey_year, season) %>% 
  summarise(cpue = mean(cpue_towspd, na.rm = T),
            percent_occurrence = mean(crab_presence, na.rm = T)) %>% 
  group_by(season) %>% 
  summarise(nyears = n(),
            avg_annual_cpue = mean(cpue, na.rm = T),
            annual_cpue_se = sd(cpue, na.rm = T) / sqrt(nyears),
            avg_percent_occurrence = mean(percent_occurrence, na.rm = T),
            po_se = sd(percent_occurrence, na.rm = T) / sqrt(nyears)) %>% 
    ungroup() %>%
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           catch_ymin         = avg_annual_cpue - (2 * annual_cpue_se),
           catch_ymax         = avg_annual_cpue + (2 * annual_cpue_se),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = avg_percent_occurrence - (2 * po_se),
           occurrence_ymax    = avg_percent_occurrence + (2 * po_se),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin),
           occurrence_ymax    = ifelse(occurrence_ymax > 1, 1, occurrence_ymax)
    )

### Seasonal Catch Plot

season_plot <- season_summary %>% 
  ggplot(aes(season, avg_annual_cpue, fill = season)) +
    geom_col() +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = NULL,
       y = "Mean Catch (C. sapidus / Hectare)") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

season_plot

### Seasonal Frequency of Occurrence

season_plot_2 <- season_summary %>% 
  ggplot(aes(season, avg_percent_occurrence, fill = season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL,
       y = "Percent Occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

season_plot_2
```

```{r}
season_plot + season_plot_2
```

>>>>

---
title: "Offshore Modeling - Manuscript Model"
author: "Adam A. Kemberling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
  code_folding: hide
editor_options: 
  chunk_output_type: console
---

@@

```{r load data}
seamap <- read_csv(str_c("~/Dropbox/SEAMAP_2019/data/offshore/", 
                        "seamap_manuscriptmods_full.csv"), 
                   col_types = cols()) %>% 
  mutate(year_f = factor(year_f),
         Season = factor(season, levels = c("Summer", "Fall")),
         region = ifelse(region == "Louisiana", "LA West", region),
         region = factor(region, levels = c("Texas", "LA West", "MS Bight", "Florida"))
)

seamap_west <- read_csv(str_c("~/Dropbox/SEAMAP_2019/data/offshore/", 
                        "seamapwest_manuscriptmods_full.csv"), 
                        col_types = cols()) %>% 
  mutate(year_f = factor(year_f),
         Season = factor(season, levels = c("Summer", "Fall")),
         region = ifelse(region == "Louisiana", "LA West", region),
         region = factor(region, levels = c("Texas", "LA West", "MS Bight", "Florida"))
)
```

@@

## Annual Catch and Occurrence {.tabset}

@@

*NOTE: Contains Data Only from Western-Gulf Subset*

```{r}
year_sums <- seamap_west %>% 
  group_by(survey_year) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(cpue_towspd)),
            `Mean Catch`      = mean(cpue_towspd, na.rm = T),
            `SD`              = sd(cpue_towspd, na.rm = T),
            `Var`             = var(cpue_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(crab_presence, na.rm = T),
            `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)
            ) %>% 
  ungroup() %>% 
    mutate(Survey_Year        = factor(survey_year),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
         )

@@

```{r, fig.height=4}

year_plot <- year_sums %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot() +
    geom_col(aes(survey_year, `Mean Catch`)) +
    geom_errorbar(aes(x = survey_year, 
                      ymin = catch_ymin,
                      ymax = catch_ymax),
                  size = .25) +
  labs(x = NULL,
       y = expression(paste("Mean CPUE (crabs ", ha^{-1},")"))
  )

@@

```{r, fig.height=4}
year_foa <- year_sums %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot() +
    geom_col(aes(survey_year, `% Occurrence`)) +
    scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
    labs(x = NULL, y = "Percent occurrence")

@@

```{r, fig.height=4}

year_plot + year_foa

## Region - Season Interactions {.tabset}

*NOTE: Contains Data Only from Western-Gulf Subset*

*NOTE 2: Mean Catch (per hectare) calculated as the average crabs/hectare from each tow within a region-season grouping. Percent occurrence is calculated separately as the average annual percent occurrence for those same groups.*

```{r }

#Mean CPUE for the regions within each season
rs_summary <- seamap_west %>%
  filter(is.na(region)      == FALSE) %>%
  group_by(region, season) %>%
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(cpue_towspd)),
            `Mean Catch`      = mean(cpue_towspd, na.rm = T),
            `SD`              = sd(cpue_towspd, na.rm = T),
            `Var`             = var(cpue_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>%
    ungroup() %>%
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           region             = factor(region,
                                       levels = c("Texas", "LA West", "MS Bight", "Florida")),
           #region             = fct_rev(region),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin)
    )

#Summary Table
rs_summary %>%
  select(Region = region, Season, Stations, `Missing Obs`, `Mean Catch`, SD, Overdispersed) %>%
  knitr::kable()
```

Average frequency of occurrence for each region and season is calculated in two steps. In the first step the frequency of occurrence for each region and for both survey seasons is calculated for every year. In the second step those annual means are averaged to get an overall measure for each region for each season, and a measure of error around them is calculated to represent the inter-annual variation in the frequency of occurrence.

@@

#Calaculate Annual average percent occurrence
rs_foa <- seamap_west %>% 
  group_by(survey_year, region, season) %>% 
  summarise(percent_occurrence = mean(crab_presence, na.rm = T)) %>% 
  group_by(region, season) %>% 
  summarise(nyears = n(),
            avg_percent_occurrence = mean(percent_occurrence, na.rm = T),
            po_se = sd(percent_occurrence, na.rm = T) / sqrt(nyears)) %>% 
    ungroup() %>%
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           region             = factor(region, levels = c("Texas", "LA West", 
                                                          "MS Bight", "Florida")),
           occurrence_ymin    = avg_percent_occurrence - (2 * po_se),
           occurrence_ymax    = avg_percent_occurrence + (2 * po_se),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin),
           occurrence_ymax    = ifelse(occurrence_ymax > 1, 1, occurrence_ymax)
    )

rs_foa %>% 
  select(Region = region, Season, `N Years` = nyears, 
         `% Occurrence` = avg_percent_occurrence, `% Occurrence SE` = po_se) %>%
  knitr::kable()

@@

```{r, fig.height=4}

# Plot
rs_plot <- rs_summary %>% 
  ggplot(aes(region, `Mean Catch`, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  position = "dodge", 
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = NULL,
       y = expression(paste("Mean CPUE (crabs ", ha^{-1},")"))) + 
  theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

```{r, fig.height=4}
rs_plot_2 <- rs_foa %>% 
  ggplot(aes(region, avg_percent_occurrence, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL,
       y = "Percent occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

```{r fig.height=4}
rs_plot + rs_plot_2

@@
